Accuracy and stability for elliptic operators in a finite volume context

### 2023-06-22 example/EBCM/hoeb_petsc/_doc/paper.tex
the Neumann domain bc equation code now compiles
getting pretty consistent solvability results

\section{Solvability of the $A$ matrix}

Overton-Katz, et al. \cite{Katz2023} present the above procedure in
the absence of boundary condition equations.   They show that, for
certain values of $P_T$ and $P_N$, the resulting $A$ matrix can have a
severe mismatch between its largest and smallest eigenvalues,
signalling that $A$ is poorly conditioned.    Their data show the
worst solvability issues occur near the domain boundary.    It stands to
reason that including boundary condition equations could improve the
conditioning of the resultant matricies. To test this, we present the
same measurement of condition number for  several values of $P_T$ and
$P_N$ where boundary condition equations are included when they are
part of the neighborhood.

>>>>

\section{Solvability of the system matrix $A$}

Overton-Katz, et al. \cite{Katz2023} present the above procedure in
the absence of boundary condition equations.   They show that, for
certain values of $P_T$ and $P_N$, the resulting $A$ matrix can have a
severe mismatch between its largest and smallest eigenvalues,
signalling that $A$ is poorly conditioned.    Their data show the
worst solvability issues occur near the domain boundary.    It stands to
reason that including boundary condition equations could improve the
conditioning of the resultant matricies. To test this, we present the
same measurement of condition number for  several values of $P_T$ and
$P_N$ where boundary condition equations are included when they are
part of the neighborhood.    We present the results for the same three
different geometries used in \cite{Katz2023}.  In figure \ref{fig::allRegSolvability} we show
the worst condition numbers where there are no cut cells.
In figure \ref{fig::unmergedSolvability}, we show the worst condition
number where we cut the unit domain with a centered sphere of radius
0.45 (this geometry includes very small volumes).    Finally, in
figure \ref{fig::mergedSolvability}, we present results all small
volumes have been merged with neighbors.

The data paint a very consistent picture,   Though adding boundary
condition equations does improve the solvability of $A$, aggressive
choices of $P_W$ and $P_T$ can lead to system matricies that cannot be
credibly solved without very high precision numerical arithmetic.
\begin{small}
\begin{table}
\begin{center}
\begin{tabular}{|ccc|ccc|} \hline
 $D$ & $P^T$  & $P^W$  & $\lambda_{max}$ & $\lambda_{min}$   & $\cali$ \\
 \hline
2 & 1 & 1 &  1.537125e+00 & 1.150772e-03 & 7.486517e-04 \\
2 & 1 & 2 &  4.940953e-01 & 9.196245e-05 & 1.861229e-04 \\
2 & 1 & 3 &  3.324177e-01 & 1.064946e-05 & 3.203639e-05 \\
2 & 1 & 4 &  2.853322e-01 & 1.705230e-06 & 5.976295e-06 \\
2 & 1 & 5 &  2.665812e-01 & 3.349335e-07 & 1.256403e-06 \\
2 & 2 & 1 &  2.407807e+00 & 1.457063e-06 & 6.051408e-07 \\
2 & 2 & 2 &  4.940956e-01 & 5.469314e-08 & 1.106934e-07 \\
2 & 2 & 3 &  3.324177e-01 & 2.879583e-09 & 8.662545e-09 \\
2 & 2 & 4 &  2.853323e-01 & 1.628939e-10 & 5.708921e-10 \\
2 & 2 & 5 &  2.665812e-01 & 1.084310e-11 & 4.067466e-11 \\
2 & 3 & 1 &  2.407807e+00 & 5.822902e-10 & 2.418342e-10 \\
2 & 3 & 2 &  9.272889e-01 & 4.520282e-11 & 4.874729e-11 \\ 
2 & 3 & 3 &  3.324177e-01 & 1.618058e-12 & 4.867544e-12 \\
2 & 3 & 4 &  2.853323e-01 & 7.227717e-14 & 2.533088e-13 \\
2 & 3 & 5 &  2.665812e-01 & 3.194386e-15 & 1.198279e-14 \\
2 & 4 & 1 &  1.880177e+00 & 1.535955e-25 & 8.169205e-26 \\
2 & 4 & 2 &  9.272889e-01 & 1.098665e-27 & 1.184814e-27 \\
2 & 4 & 3 &  6.561304e-01 & 5.104096e-28 & 7.779089e-28 \\
2 & 4 & 4 &  2.868930e-01 & 3.731928e-27 & 1.300808e-26 \\
2 & 4 & 5 &  5.327046e-01 & 5.118416e-25 & 9.608357e-25 \\
\hline
3 & 1 & 1 &  4.639099e+00 & 3.887142e-03 & 8.379088e-04 \\ 
3 & 1 & 2 &  1.079504e+00 & 2.369894e-04 & 2.195354e-04 \\ 
3 & 1 & 3 &  6.747118e-01 & 2.359590e-05 & 3.497182e-05 \\ 
3 & 1 & 4 &  5.721834e-01 & 3.540976e-06 & 6.188534e-06 \\ 
3 & 1 & 5 &  5.333668e-01 & 6.792558e-07 & 1.273525e-06 \\ 
3 & 2 & 1 &  4.639156e+00 & 3.176883e-06 & 6.847975e-07 \\ 
3 & 2 & 2 &  1.079505e+00 & 1.512048e-07 & 1.400686e-07 \\ 
3 & 2 & 3 &  6.747119e-01 & 6.991076e-09 & 1.036157e-08 \\ 
3 & 2 & 4 &  5.721835e-01 & 3.641527e-10 & 6.364265e-10 \\ 
3 & 2 & 5 &  2.669165e-01 & 1.146071e-11 & 4.293744e-11 \\ 
3 & 3 & 1 &  4.639157e+00 & 1.340233e-09 & 2.888959e-10 \\ 
3 & 3 & 2 &  1.079505e+00 & 7.151353e-11 & 6.624659e-11 \\ 
3 & 3 & 3 &  6.747119e-01 & 3.977531e-12 & 5.895155e-12 \\ 
3 & 3 & 4 &  2.869943e-01 & 5.757645e-14 & 2.006188e-13 \\ 
3 & 3 & 5 &  2.669165e-01 & 1.723124e-15 & 6.455665e-15 \\ 
3 & 4 & 1 &  6.402997e+00 & 5.729285e-26 & 8.947817e-27 \\ 
3 & 4 & 2 &  1.079505e+00 & 1.226846e-27 & 1.136489e-27 \\ 
3 & 4 & 3 &  6.907344e-01 & 1.049444e-28 & 1.519317e-28 \\ 
3 & 4 & 4 &  2.903904e-01 & 8.442770e-26 & 2.907386e-25 \\ 
3 & 4 & 5 &  5.333668e-01 & 5.987514e-25 & 1.122588e-24 \\ 
\hline
\end{tabular}
\end{center}
\label{fig::allRegSolvability}
\caption
    {
      We show the lowest value of
      $\cali$ for an entire Cartesian grid of size $32^D$.
      We vary Taylor power $P^T$, weighting
      function exponents $P^W$, and dimensionality $D$. 
    }
\end{table}
\end{small}

\begin{small}
\begin{table}
\begin{center}
\begin{tabular}{|ccc|ccc|} \hline
 $D$ & $P^T$  & $P^W$  & $\lambda_{max}$ & $\lambda_{min}$   & $\cali$  \\
   \hline
2 & 1 & 1 &  3.244145e+00 & 5.576001e-03 & 1.718789e-03 \\ 
2 & 1 & 2 &  1.108699e+00 & 5.218520e-04 & 4.706884e-04 \\ 
2 & 1 & 3 &  6.816993e-01 & 4.992473e-05 & 7.323570e-05 \\ 
2 & 1 & 4 &  5.736425e-01 & 7.273424e-06 & 1.267937e-05 \\ 
2 & 1 & 5 &  5.336857e-01 & 1.374234e-06 & 2.574988e-06 \\ 
2 & 2 & 1 &  2.762140e+00 & 6.863083e-06 & 2.484698e-06 \\ 
2 & 2 & 2 &  9.453294e-01 & 7.653732e-07 & 8.096365e-07 \\ 
2 & 2 & 3 &  6.816994e-01 & 4.112367e-08 & 6.032523e-08 \\ 
2 & 2 & 4 &  5.736426e-01 & 3.156232e-09 & 5.502089e-09 \\ 
2 & 2 & 5 &  5.336858e-01 & 4.220417e-10 & 7.908057e-10 \\ 
2 & 3 & 1 &  2.762140e+00 & 2.066812e-09 & 7.482649e-10 \\ 
2 & 3 & 2 &  9.453294e-01 & 1.559241e-10 & 1.649416e-10 \\ 
2 & 3 & 3 &  6.574277e-01 & 2.360967e-11 & 3.591220e-11 \\ 
2 & 3 & 4 &  5.705795e-01 & 4.724557e-12 & 8.280278e-12 \\ 
2 & 3 & 5 &  5.336858e-01 & 2.690617e-13 & 5.041575e-13 \\ 
2 & 4 & 1 &  2.762140e+00 & 3.535738e-13 & 1.280072e-13 \\ 
2 & 4 & 2 &  9.453294e-01 & 3.267022e-14 & 3.455962e-14 \\ 
2 & 4 & 3 &  6.574277e-01 & 3.738186e-15 & 5.686079e-15 \\ 
2 & 4 & 4 &  5.691873e-01 & 3.830503e-16 & 6.729777e-16 \\ 
2 & 4 & 5 &  5.328410e-01 & 8.531691e-17 & 1.601170e-16 \\
\hline
3 & 1 & 1 &  7.242861e+00 & 4.674236e-03 & 6.453577e-04 \\ 
3 & 1 & 2 &  1.181250e+00 & 2.657202e-04 & 2.249483e-04 \\ 
3 & 1 & 3 &  6.863432e-01 & 3.329115e-05 & 4.850511e-05 \\ 
3 & 1 & 4 &  5.741086e-01 & 4.562521e-06 & 7.947139e-06 \\ 
3 & 1 & 5 &  5.350591e-01 & 1.455556e-06 & 2.720366e-06 \\ 
3 & 2 & 1 &  7.433403e+00 & 2.264150e-06 & 3.045914e-07 \\ 
3 & 2 & 2 &  1.181248e+00 & 1.685503e-07 & 1.426883e-07 \\ 
3 & 2 & 3 &  6.804738e-01 & 1.353871e-08 & 1.989600e-08 \\ 
3 & 2 & 4 &  5.806933e-01 & 2.036542e-09 & 3.507086e-09 \\ 
3 & 2 & 5 &  5.350591e-01 & 1.066432e-10 & 1.993110e-10 \\ 
3 & 3 & 1 &  7.433404e+00 & 8.925217e-10 & 1.200690e-10 \\ 
3 & 3 & 2 &  1.182547e+00 & 8.635683e-11 & 7.302615e-11 \\ 
3 & 3 & 3 &  6.804738e-01 & 4.819403e-12 & 7.082423e-12 \\ 
3 & 3 & 4 &  5.729018e-01 & 3.780488e-13 & 6.598842e-13 \\ 
3 & 3 & 5 &  5.350591e-01 & 6.218451e-14 & 1.162199e-13 \\ 
3 & 4 & 1 &  7.433404e+00 & 1.727339e-13 & 2.323753e-14 \\ 
3 & 4 & 2 &  1.182547e+00 & 2.186136e-14 & 1.848668e-14 \\ 
3 & 4 & 3 &  6.804738e-01 & 1.827956e-15 & 2.686299e-15 \\ 
3 & 4 & 4 &  5.728477e-01 & 7.990908e-17 & 1.394944e-16 \\ 
3 & 4 & 5 &  5.338853e-01 & 1.743421e-17 & 3.265535e-17 \\ 
\hline
\end{tabular}
\end{center}
\label{fig::unmergedSolvability}
\caption
    {
      We show the lowest value of
      $\cali$ for an entire Cartesian grid of size $32^D$.
      The grid covers the unit square and is cut by a sphere of radius 0.45.
      We vary Taylor power $P^T$, weighting
      function exponents $P^W$, and dimensionality $D$.    This grid
      includes very small volumes.
    }
\end{table}
\end{small}

\begin{small}
\begin{table}
\begin{center}
  \begin{tabular}{|ccc|ccc|} \hline
 $D$ & $P^T$  & $P^W$  & $\lambda_{max}$ & $\lambda_{min}$   & $\cali$ \\
  \hline
2 & 1 & 1 &  3.110810e+00 & 6.876828e-04 & 2.210623e-04 \\ 
2 & 1 & 2 &  1.820276e+00 & 2.935551e-04 & 1.612695e-04 \\ 
2 & 1 & 3 &  6.689318e-01 & 2.783647e-05 & 4.161332e-05 \\ 
2 & 1 & 4 &  3.112964e-01 & 3.804498e-06 & 1.222146e-05 \\ 
2 & 1 & 5 &  2.931197e-01 & 4.648824e-07 & 1.585981e-06 \\ 
2 & 2 & 1 &  4.649575e+00 & 1.222691e-06 & 2.629683e-07 \\ 
2 & 2 & 2 &  2.800875e+00 & 4.389241e-07 & 1.567096e-07 \\ 
2 & 2 & 3 &  1.156823e+00 & 4.756446e-08 & 4.111647e-08 \\ 
2 & 2 & 4 &  1.020254e+00 & 3.829097e-09 & 3.753083e-09 \\   
2 & 2 & 5 &  9.591803e-01 & 4.980114e-10 & 5.192052e-10 \\ 
2 & 3 & 1 &  8.598864e-01 & 3.926204e-10 & 4.565956e-10 \\ 
2 & 3 & 2 &  3.682481e+00 & 6.386263e-10 & 1.734228e-10 \\ 
2 & 3 & 3 &  1.036128e+00 & 1.177339e-11 & 1.136288e-11 \\ 
2 & 3 & 4 &  9.070960e-01 & 5.028483e-13 & 5.543495e-13 \\ 
2 & 3 & 5 &  8.510270e-01 & 3.323827e-14 & 3.905666e-14 \\ 
2 & 4 & 1 &  9.197999e-01 & 3.623068e-14 & 3.938974e-14 \\ 
2 & 4 & 2 &  7.945940e-02 & 1.618085e-15 & 2.036366e-14 \\ 
2 & 4 & 3 &  1.186857e+00 & 4.379402e-15 & 3.689917e-15 \\ 
2 & 4 & 4 &  1.039000e+00 & 3.390648e-16 & 3.263375e-16 \\ 
2 & 4 & 5 &  2.958494e-01 & 1.292309e-17 & 4.368131e-17 \\
\hline
3 & 1 & 1 &  5.044398e+00 & 6.813020e-04 & 1.350611e-04 \\
3 & 1 & 2 &  2.141787e+00 & 2.336010e-05 & 1.090683e-05 \\
3 & 1 & 3 &  1.663929e+00 & 8.355671e-07 & 5.021651e-07 \\
3 & 1 & 4 &  1.479289e+00 & 9.414478e-08 & 6.364192e-08 \\
3 & 1 & 5 &  3.223939e-01 & 1.193480e-08 & 3.701931e-08 \\
3 & 2 & 1 &  6.441253e+00 & 1.723363e-07 & 2.675509e-08 \\
3 & 2 & 2 &  3.588187e+00 & 1.066316e-08 & 2.971741e-09 \\
3 & 2 & 3 &  2.800032e+00 & 9.932906e-10 & 3.547425e-10 \\
3 & 2 & 4 &  2.491816e+00 & 2.063069e-10 & 8.279378e-11 \\
3 & 2 & 5 &  6.820526e-01 & 2.306368e-11 & 3.381511e-11 \\
3 & 3 & 1 &  2.508964e+00 & 1.541288e-11 & 6.143127e-12 \\
3 & 3 & 2 &  4.951047e+00 & 1.076120e-11 & 2.173521e-12 \\
3 & 3 & 3 &  3.867831e+00 & 1.354407e-12 & 3.501722e-13 \\
3 & 3 & 4 &  2.166936e+00 & 8.701202e-14 & 4.015441e-14 \\
3 & 3 & 5 &  2.040600e+00 & 3.890155e-15 & 1.906378e-15 \\
3 & 4 & 1 &  2.418622e+00 & 1.859733e-15 & 7.689226e-16 \\
3 & 4 & 2 &  1.286816e+00 & 1.115219e-15 & 8.666503e-16 \\
3 & 4 & 3 &  9.013057e-01 & 9.738926e-17 & 1.080535e-16 \\
3 & 4 & 4 &  7.856540e-01 & 2.968168e-17 & 3.777958e-17 \\
3 & 4 & 5 &  9.668729e-01 & 1.139377e-17 & 1.178415e-17 \\
\hline
\end{tabular}
\end{center}
\label{fig::mergedSolvability}
\caption
    {
      We show the lowest value of
      $\cali$ for an entire Cartesian grid of size $32^D$.
      The grid covers the unit square and is cut by a sphere of radius 0.45.
      We vary Taylor power $P^T$, weighting
      function exponents $P^W$, and dimensionality $D$.    This grid
      has had its small volumes merged with neighboring volumes.
    }
\end{table}
\end{small}
